---
title: "Take-home_Ex02"
author: "LIANG YAO"
date: '27 May 2023'
date-modified: "`r Sys.Date()`"
execute: 
  warning: false
  echo: true
  eval: true
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
---

## 1. Task and Questions:

### Objectives:

The country of Oceanus has sought FishEye International's help in identifying companies possibly engaged in illegal, unreported, and unregulated (IUU) fishing. As part of the collaboration, FishEye's analysts received import/export data for Oceanus' marine and fishing industries. However, Oceanus has informed FishEye that the data is incomplete. To facilitate their analysis, FishEye transformed the trade data into a knowledge graph. Using this knowledge graph, they hope to understand business relationships, including finding links that will help them stop IUU fishing and protect marine species that are affected by it. FishEye analysts found that node-link diagrams gave them a good high-level overview of the knowledge graph. However, they are now looking for visualizations that provide more detail about patterns for entities in the knowledge graph. There are two main parts to this analysis.

First, FishEye knows from past experience that companies caught fishing illegally will shut down but will then often start up again under a different name. FishEye wants your help to visualize temporal patterns so they can compare the activities of companies over time to determine if the companies have returned to their nefarious acts.

Second, FishEye has been using several tools, including artificial intelligence, to reason on the knowledge graph and suggest links that could extend the dataset. They have supplied 12 groups of link suggestions and need your help evaluating these groups to identify which tools are most reliable for completing the graph. FishEye is especially interested in identifying new temporal patterns or anomalies that are only present when new links are added.

### Questions:

1.  Use visual analytics to identify temporal patterns for individual entities and between entities in the knowledge graph FishEye created from trade records. Categorize the types of business relationship patterns you find. Limit your response to 600 words and 6 images.

2.  Evaluate the sets of predicted knowledge graph links FishEye has provided using visual analytics. Which sets are most reliable for completing the graph? Limit your response to 600 words and 6 images.

3.  Illustrate how your visual analytics approach can be used to identify new patterns and/or anomalies that are present in the knowledge graph after you have added the links you deemed reliable in question 2. Limit your response to 300 words and 4 images.

4.  Identify companies that fit a pattern of illegal fishing. Use visualizations to support your conclusions and your confidence in them. Limit your response to 300 words and 4 images.

5.  Reflection: What was the most difficult aspect of working with this knowledge graph? Did you have the tools and resources you needed to complete the challenge? What additional resources would have helped you? Limit your response to 300 words

## 2. Load packages and data:

The data should be processed by using appropriate **tidyverse** family of packages and the statistical graphics must be prepared using ggplot2 and its extensions.

```{r}
pacman::p_load(ggstatsplot, tidyverse, ggplot2, plotly, ggmosaic, ggridges, gganimate,ggdist,jsonlite)

mc2 <- jsonlite::fromJSON("data/mc2_challenge_graph.json")
```

## 3. Data checking/cleaning:

### 3.1 Find the nodes and edges:

```{r}
#view(mc2[["nodes"]])
mc2_nodes <- as_tibble(mc2$nodes) %>%
  select(id, shpcountry, rcvcountry)
```

```{r}
#view(mc2[["links"]])
mc2_edges <- as_tibble(mc2$links) %>%
  select(source, target, weightkg, hscode, arrivaldate, volumeteu, valueofgoods_omu, valueofgoodsusd)
```

### 3.2 Data wranggling

### 3.2.1 Check and deal with NA:

-   It's confirmed from the summary table that "wages" and "rent adjustment" got only positive numbers and other categories are negative numbers.

```{r}

```

::: callout-warning
Here is warning message.
:::

### 3.2.2 Aggregate by links?:

-   Distribution of resources:

    ```{r}

    ```

-   Distribution of taragets:

    ```{r}

    ```

### 3.3 Combine information into one table

-   .

```{r}

```

::: callout-note
Here is a note
:::

-   Then summarize into a new table named "fi_ave" in which transpose the monthly average Income/Expenses categories into column names;

```{r}

```

-   Combine "fi_ave" with "parti" into a new table named "joy".

```{r}

```

# 4. Analysis & Visualization

## 4.1 Visualizing original links

### 4.1.1 ?

From the chart below we can find that different interest group attract people with quite different education level.

```{r}

```

### 4.1.2 ?

```{r}


```

## 4.2 Finding pattern

### 4.2.1 new patterns?

Citizens' monthly income apparently goes up with their educational level, indicating local labor market do care about employees' education background.

```{r}

```

### 4.2.2 Visualizing

Monthly income distributing quite evenly over different household size and joviality level.

```{r}

```
