---
title: "Take-home_Ex01"
author: "LIANG YAO"
date: '13 May 2023'
date-modified: "`r Sys.Date()`"
execute: 
  warning: false
  echo: true
  eval: true
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
---

## 1. Task and data:

### Objectives:

City of Engagement, with a total population of 50,000, is a small city located at Country of Nowhere. The city serves as a service centre of an agriculture region surrounding the city. The main agriculture of the region is fruit farms and vineyards. The local council of the city is in the process of preparing the Local Plan 2023. A sample survey of 1000 representative residents had been conducted to collect data related to their household demographic and spending patterns, among other things. The city aims to use the data to assist with their major community revitalization efforts, including how to allocate a very large city renewal grant they have recently received.

### Data

For the purpose of this study, two data sets are provided. They are:

*Participants.csv*

Contains information about the residents of City of Engagement that have agreed to participate in this study.

-   participantId (integer): unique ID assigned to each participant.

-   householdSize (integer): the number of people in the participant's household

-   haveKids (boolean): whether there are children living in the participant's household.

-   age (integer): participant's age in years at the start of the study.

-   educationLevel (string factor): the participant's education level, one of: {"Low", "HighSchoolOrCollege", "Bachelors", "Graduate"}

-   interestGroup (char): a char representing the participant's stated primary interest group, one of {"A", "B", "C", "D", "E", "F", "G", "H", "I", "J"}. Note: specific topics of interest have been redacted to avoid bias.

-   joviality (float): a value ranging from \[0,1\] indicating the participant's overall happiness level at the start of the study.

*FinancialJournal.csv*

Contains information about financial transactions.

-   participantId (integer): unique ID corresponding to the participant affected

-   timestamp (datetime): the time when the check-in was logged

-   category (string factor): a string describing the expense category, one of {"Education", "Food", "Recreation", "RentAdjustment", "Shelter", "Wage"}

-   amount (double): the amount of the transaction

For explanation of Rent Adjustment, please refer to this [link](https://www.lawinsider.com/dictionary/rent-adjustment-amount).

## 2. Load packages and data:

The data should be processed by using appropriate **tidyverse** family of packages and the statistical graphics must be prepared using ggplot2 and its extensions.

```{r}
pacman::p_load(ggstatsplot, tidyverse, ggplot2, plotly, ggmosaic, ggridges, gganimate,ggdist)

fi <- read.csv("data/FinancialJournal.csv")
parti <- read.csv("data/Participants.csv")
```

## 3. Data checking/cleaning:

### 3.1 Check the summary of tables to get a whole picture of the data:

```{r}
summary(fi)
summary(parti)
```

### 3.2 Some further check on each table

### 3.2.1 Further check of fi table by summarizing by categories:

-   It's confirmed from the summary table that "wages" and "rent adjustment" got only positive numbers and other categories are negative numbers.

```{r}
fi$month <- months.Date(strptime(fi$timestamp, "%Y-%m-%d"))
fi$year <- format(strptime(fi$timestamp, "%Y-%m-%d"), format="%Y")
fi$yearmonth <- paste(fi$year,fi$month)

fi %>%
  group_by(year, category) %>%
  summarise(median = median(amount),
            mean = mean(amount),
            min = min(amount),
            max = max(amount),
            n = n(),
            neg = sum(amount<0))
```

::: callout-warning
Here the Rent Adjustment amount which have much less N of records compare to other categories need to be carefully treated.
:::

### 3.2.2 Further check of parti table:

-   Distribution of joviality:

    ```{r}
    d <- highlight_key(parti)
    p <-ggplot(data=parti, aes(x = joviality)) +
      geom_histogram(bins=10, 
                     boundary = 100,
                     color="white", 
                     fill="#4575B4") +
      theme_light()+
      ggtitle("Distribution of joviality")

    highlight(ggplotly(p),"plotly_selected")  
    ```

-   Distribution of other variables:

    ```{r}
    d <- highlight_key(parti)
    p1 <- ggplot(data=d, aes(x = householdSize)) +
      geom_bar(stat = "count", 
               boundary = 100,
               color="white", 
               fill="#4575B4") +
      xlab("Household Size")

    p2 <- ggplot(data=d, aes(x = age)) +
      geom_bar(stat = "count", 
               boundary = 100,
               color="white", 
               fill="#ABD9E9") +
      xlab("Age")

    p3 <- ggplot(data=d, aes(x = educationLevel)) +
      geom_bar(stat = "count", 
               boundary = 100,
               color="white", 
               fill="#FEE090") +
      xlab("Education Level")+
      theme(axis.text.x = element_text(angle = 30))

    p4 <- ggplot(data=d, aes(x = interestGroup)) +
      geom_bar(stat = "count", 
               boundary = 100,
               color="white", 
               fill="#F46D43") +
      xlab("Interest Group")

    subplot(ggplotly(p1),
            ggplotly(p2),
            ggplotly(p3),
            ggplotly(p4),
            nrows = 2,
            titleX = TRUE,
            margin = 0.1)
    ```

### 3.3 Combine information into one table

-   Find monthly average income and expenses of each participant ID;

-   Re-categorize wage as "Income", and all others together as "Expenses";

-   And for the convenience of latter using, I will create a column named "absave" to save the absolute number of the amount.

```{r}
fi_ave_by_cat <- fi%>%
  group_by(participantId, category) %>%
  summarise(total_amount = sum(amount),
            n_months = length(unique(yearmonth)))
fi_ave_by_cat$ave = ifelse(fi_ave_by_cat$category == 'RentAdjustment', 
                           fi_ave_by_cat$total_amount/12, 
                           fi_ave_by_cat$total_amount/fi_ave_by_cat$n_months)

fi_ave_by_cat$bicat <- ifelse(fi_ave_by_cat$category == 'Wage','Income', 'Expenses')
fi_ave_by_cat$absave <- abs(fi_ave_by_cat$ave)
```

::: callout-note
Here I divided the amount of "Rent Adjustment" by 12 rather than the unique months of it's record, that's because this amount is an annual amount occurred only once in a year.
:::

-   Then summarize into a new table named "fi_ave" in which transpose the monthly average Income/Expenses categories into column names;

```{r}
fi_ave <- fi_ave_by_cat%>%
  group_by(participantId, bicat) %>%
  summarise(mon_ave = abs(sum(ave))) %>%
  pivot_wider(names_from = bicat,values_from = mon_ave, names_prefix = "mon_ave_")
```

-   Combine "fi_ave" with "parti" into a new table named "joy".

```{r}
joy <- full_join(fi_ave,parti,by="participantId")
summary(joy)
```

# 4. Analysis & Visualization

## 4.1 Finding demographic characteristics of the city

### 4.1.1 How interest group varies by education levels?

From the chart below we can find that different interest group attract people with quite different education level.

```{r}
d <- highlight_key(joy)
p <- ggplot(data = joy) +
  geom_mosaic(aes(x = product(interestGroup), fill= educationLevel)) +
  theme_mosaic() +
  scale_fill_manual(values = c("#4575B4", "#ABD9E9", "#FEE090", "#F46D43"))+
  xlab("")+
  ylab("")

highlight(ggplotly(p),"plotly_selected")  
```

### 4.1.2 How interest group changes with people's age?

```{r}
ggplot(joy, aes(x = age, y = interestGroup, fill = interestGroup)) +
  geom_density_ridges(alpha=0.8, color = "white") +
  theme_ridges() + 
  theme(legend.position = "none",
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 8))+
  scale_fill_manual(values = c("#F4EDCA","#E7B800", "#C4961A",  "#D16103", "#C3D7A4", "#52854C", "#4E84C4", "#ABD9E9", "#FEE090", "#F46D43")) +
  xlab("Age") +
  ylab("Interest Groups")

```

## 4.2 Finding financial characteristics of the city

### 4.2.1 How monthly average income varies between different education level groups

Citizens' monthly income apparently goes up with their educational level, indicating local labor market do care about employees' education background.

```{r}
ggbetweenstats(
  data = joy,
  x = educationLevel, 
  y = mon_ave_Income,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

### 4.2.2 Looks into monthly average income's distribution together with other demographic information

Monthly income distributing quite evenly over different household size and joviality level.

```{r}
d <- highlight_key(joy)
p <- ggplot(joy, aes(x = mon_ave_Income, y = joviality, 
                      size = householdSize, 
                      colour = haveKids)) +
  geom_point(alpha = 0.6) +
  scale_colour_manual(values = c("#4E84C4", "#FEE090")) +
  scale_size(range = c(1, 3))+
  labs(title = 'Do people earn more feel happier?', 
       x = 'Monthly average income', 
       y = 'Joviality')

highlight(ggplotly(p),"plotly_selected")  
```

### 4.2.3 How monthly expenses varies between different household size

The amount of monthly expenses of single household is quite different from other household size.

```{r}
ggbetweenstats(
  data = joy,
  x = householdSize, 
  y = mon_ave_Expenses,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

### 4.2.4 Looks into monthly average expenses by categories

And people's most big amount of expense seems to be shelter expenses.

```{r}
fi_ave_by_cat %>%
  filter(bicat != "Income" & category != "RentAdjustment") %>%
  ggplot(aes(x = category, y = absave)) +
  stat_pointinterval(.width = 0.95,
  .point = median,
  .interval = qi) +
  theme_light() +
  labs(
    title = "Confidence intervals of mean expense by category",
    subtitle = "Mean Point + Multiple-interval plot")
```

## 4.3 Finding correlation between characteristics and joviality

### 4.3.1 Do people earn more feel happier?

No matter we look into the correlation between joviality and monthly income as a whole or by educational level group, the correlation is relatively week.

```{r}
ggscatterstats(data = joy,
                     x = mon_ave_Income,
                     y = joviality,
                     marginal = FALSE,
                     point.args = list(size = 1, alpha = 0.6, stroke = 0),
               title = "Joviality negatively correlated with monthly average income")
```

```{r}
grouped_ggscatterstats(
  data = joy,
  x = mon_ave_Income,
  y = joviality,
  marginal = FALSE,
  point.args = list(size = 1, alpha = 0.6, stroke = 0),
  grouping.var = educationLevel,
  type = "r",
  annotation.args = list(title = "Relationship between joviality and income by education level"),
  plotgrid.args = list(nrow = 2, ncol = 2)
)
```

### 4.3.2 Or do people spend more feel happier?

-   To look at the correlation between joviality and monthly average expenses, seemingly there is only a weakly positive correlation.

```{r}
ggscatterstats(data = joy,
                    x = mon_ave_Expenses,
                    y = joviality,
                    marginal = FALSE,
                    point.args = list(size = 1, alpha = 0.6, stroke = 0),
               title = "Joviality positively correlated with monthly average expenses")
```

-   Funny part is, when we look at this correlation by different household size, the correlation between joviality and monthly average expenses is much stronger for size 1 households, indicating single persons enjoy spending money!

```{r}
grouped_ggscatterstats(
  ## arguments relevant for ggscatterstats
  data = joy,
  x = mon_ave_Expenses,
  y = joviality,
  marginal = FALSE,
  point.args = list(size = 1, alpha = 0.6, stroke = 0),
  grouping.var = householdSize,
  type = "r", 
  annotation.args = list(title = "Relationship between joviality and expenses by household size"),
  plotgrid.args = list(nrow = 1)
)
```

-   Then let's dig deeper to see how single person's spend their money.

    Unlike what we have seen in analysis of whole data set which indicating people spend the most on shelter, the average amount single citizens spend on food and recreation comes much closer to the average amount spend on shelter.

    ```{r}
    singleId <- joy %>%
      filter(householdSize == 1) %>%
      subset(select = "participantId")

    expense_cat <- fi_ave_by_cat%>%
      filter(bicat == "Expenses" & category != "RentAdjustment") %>%
      subset(select = c("participantId", "category", "absave")) %>%
      group_by(participantId, category) %>%
      summarise(absave = sum(absave))

    single <- left_join(singleId, expense_cat, by = "participantId")

    single %>%
      ggplot(aes(x = category, y = absave)) +
      stat_pointinterval(.width = 0.95,
      .point = median,
      .interval = qi) +
      theme_light() +
      labs(
        title = "Confidence intervals of single citizens' mean expense by category",
        subtitle = "Mean Point + Multiple-interval plot")
    ```

-   If this is the case, do person's joviality correlated with their spending on recreation?

    Yes, they do!

    ```{r}
    joy_withcat <- left_join(joy, 
                     expense_cat %>% pivot_wider(names_from = category, names_prefix = "mon_ave_", values_from = absave),
                     by = "participantId")

    ggscatterstats(data = joy_withcat,
                        x = mon_ave_Recreation,
                        y = joviality,
                        marginal = FALSE,
                        point.args = list(size = 1, alpha = 0.6, stroke = 0),
                   title = "Joviality positively correlated with monthly average spending on recreation")
    ```

-   How about expenses on food?

    No, they don't.

    ```{r}
    ggscatterstats(data = joy_withcat,
                        x = mon_ave_Food,
                        y = joviality,
                        marginal = FALSE,
                        point.args = list(size = 1, alpha = 0.6, stroke = 0),
                   title = "Joviality positively correlated with monthly average spending on recreation")
    ```
